Peer Assessment 1
========================================================

## Load and Preprocess the data

```{r, echo=TRUE}
activity <- read.csv("activity.csv", h=T)
activity["weekday"] <- weekdays(as.Date(activity$date))
activity["weekend"] <- with(activity, ifelse(weekday == "Sunday" | weekday == "Saturday", "Weekend", "Weekday"))
```


```{r, echo=TRUE}
mean_act <- mean(activity$steps, na.rm = TRUE)
median_act <- median(activity$steps, na.rm = TRUE)
```

The **mean** is **`r mean_act`** and **median** is **`r median_act`**

## Histogram of steps and mean and median. Median is red and mean is blue

```{r, echo=TRUE}
hist(activity$steps, main = "Histogram of Steps", xlab = "Steps")
abline(v=mean_act, col = "blue", lwd = 2)
abline(v=median_act, col = "red", lwd = 2)
```

## The average daily activity pattern

```{r, echo=TRUE}
library(plyr)
mean_step_df <- ddply(activity, .(interval), summarise, mean_steps = mean(steps, na.rm=TRUE))
plot(mean_step_df$mean_steps ~ mean_step_df$interval, type = "l", ylab = "Avg number of steps", xlab = "Interval", main = "Average number of steps vs interval")
```



```{r, echo=TRUE}
max_int <- mean_step_df[mean_step_df$mean_steps == max(mean_step_df$mean_steps), "interval"]
```

The 5 minute interval with the most steps in the day is `r max_int`

## Inputting missing values

### The number of missing values

```{r, echo=TRUE}
num_nas <- sum(is.na(activity$steps))
new_data <- activity
for(i in mean_step_df$interval){
  new_data[is.na(new_data$steps) & new_data$interval == i,"steps"] <- mean_step_df[mean_step_df$interval == i, "mean_steps"]
  }
mean_act_new <- mean(new_data$steps)
median_act_new <- median(new_data$steps)
```

The __mean__ for the new data set is **`r mean_act_new`** and __median__ is **`r median_act_new`**

```{r, echo=TRUE}
hist(new_data$steps, main = "Histogram of new data Steps")
abline(v=mean_act_new, col = "blue", lwd = 2)
abline(v=median_act_new, col = "red", lwd = 2)
```

## Differences between weekday and weekend

```{r, echo=TRUE}
library(lattice)
mean_step_weekend_df <- ddply(activity, .(interval, weekend), summarise, mean_steps = mean(steps, na.rm=TRUE))
xyplot(mean_step_weekend_df$mean_steps ~ mean_step_weekend_df$interval| mean_step_weekend_df$weekend, type = "l", ylab = "Avg number of steps", xlab = "Interval", main = "Average number of steps vs interval ordered by weekend or weekday", layout = c(1,2))
```